My First Markdown File
======================

This is my first R markdown file.

Here, we're going to load some data.

```{r}
library(datasets)
data(airquality)
summary(airquality)
```

Let's first make a pairs plot of the data.

```{r}
pairs(airquality)
```

Here`s a regression model of ozone on wind, solar radiation and temperature.

```{r}
library(stats)
fit <- lm(Ozone ~ Wind + Solar.R + Temp, data=airquality)
summary(fit)
```

Here's an unordered list:

* First element

* Second Element

```{r simulation, echo=FALSE, results='hide'}
set.seed(1)
x <- rnorm(100)
mean(x)
```

```{r computetime, echo=FALSE}
time <- format(Sys.time(), "%a %b %d %X %Y")
rand <- rnorm(1)
```

The current time is `r time`. My favourite random number is `r rand`.

Let's simulate some data.
```{r simulatedata, echo=TRUE}
x <- rnorm(100); y <- x + rnorm(100, sd=0.5)
```

Here's the scatterplot of the data:

```{r scatterplot, fig.height=4}
par(mar = c(5,4,1,1), las = 1)
plot(x, y, main="My simulate Data")
```

```{r fitmodel}
library(datasets)
data(airquality)
fit <- lm(Ozone ~ Wind+Temp+Solar.R, data=airquality)
```

Here is a table of regression coefficients.

```{r showtable, results="asis"}
library(xtable)
xt <- xtable(summary(fit))
print(xt, type="html")
```