Reproducible Research Assignment 1 - Activity Monitoring Data
========================================
 
### The objective of theis assignment is to understand the patterns of a activity monitor (specifically a step monitor) to gain knowledge of how is the usage of this monitor for a single subject.

###Firstly, we load the data and explore how the main features are allocated:

```{r}
setwd('D:/Coursera/datasciencecoursera/Reproducible Research/Assignment 1')
data <- read.csv('./Data/activity.csv')
head(data)
summary(data)
str(data)
```

### After, we load the libraries used for data manipulation:

```{r}
library(plyr)
library(dplyr)
library(tidyr)

```

### Grouping the data by date and plotting the histogram for the steps by day

```{r}
steps <- data %>%
    group_by(date) %>%
    summarise_all(sum)
head(steps)
hist(steps$steps, breaks =15)
```

### Calculating the average and median steps per day. Since the null and zero values influences on the result, I've opted to remove them for a clearer output.

```{r}
medianSteps <- data %>%
    filter(steps != 0) %>%
    group_by(date) %>%
    summarise(sum = sum(steps), mean=mean(steps), median=median(steps))
medianSteps
```

### Now we group the date by interval to understand how many steps the subject takes along the day:

```{r}
activityPattern <- data %>%
    filter(steps !=0) %>%
    group_by(interval) %>%
    summarise(steps=mean(steps))
activityPattern
```

### Plotting a time series of the activity pattern:

```{r}
plot(activityPattern$interval,activityPattern$steps,type="l",xlab="Interval",ylab="Date")
```

### Checking how many steps on average the subject has taken (zeroes values are ignored here)

```{r}
sum(activityPattern$steps)
```

### Checking the interval of the day where the subject is most active:

```{r}
activityPattern[which.max(activityPattern$steps),]
```

### Count the NA values and how much it representes in percent

```{r}
sum(is.na(data$steps))
sum(is.na(data$steps))/(sum(is.na(data$steps))+sum(!is.na(data$steps)))*100
```

### Since the missing data represents 13.11% of the data, it is suitable to fill the missing values. The method chosen is to replace the missing values by the average of that interval

### Firstly, we recalculate the activity pattern, replacing the NA values by 0

```{r}
activityPattern2 <- data %>%
    replace(., is.na(.), 0) %>%
    group_by(interval) %>%
    summarise(steps=mean(steps))
activityPattern2
```


### Now we replace the missing data:

```{r}
fullData <- data %>%
    rename(steps.x = steps) %>%
    join(activityPattern2, by="interval") %>%
    mutate(steps.x = ifelse(is.na(steps.x), steps,steps.x)) %>%
    select(steps=steps.x, date, interval)
head(fullData)
```


### Check if there is still some data missing:

```{r}
sum(is.na(fullData$steps))
```

### Replotting the histogram of the average steps per day:

```{r}
fullSteps <- fullData %>%
    group_by(date) %>%
    summarise_all(sum)
head(fullSteps)
hist(fullSteps$steps, breaks =15)
```

### We notice that there is an increase of the average steps close to 10k compared with the histogram with missing data.

```{r}
hist(steps$steps, breaks =15)
```

### Recalculating the average and median steps per day:

```{r}
fullMedianSteps <- fullData %>%
    filter(steps !=0) %>%
    group_by(date) %>%
    summarise(sum = sum(steps), mean=mean(steps), median=median(steps, na.rm=T))
fullMedianSteps
```

### Now we separate the weekdays from the weekend, so we can see if there is a change of pattern. Firstly, we create a column with the weekday as string, after we create the dummy variables for weekend (1 if TRUE).

OBS: Super fast portuguese language course: sábado = saturday and domingo = sunday ;)

```{r}
fullData2 <- fullData %>%
    mutate(weekday = weekdays(as.Date(fullData$date))) %>%
    mutate(weekend = ifelse(weekday == "sábado" | weekday =="domingo", 1, 0))
head(fullData2)
```

### Checking if the dummy variable worked

```{r}
head(fullData2[fullData2$date == "2012-10-06", ],10)
```

### And lastly, we filter the data by weekend and plot the results: 

```{r}
weekendPlot <- fullData2 %>%
    filter(weekend==1) %>%
    group_by(interval) %>%
    summarise(steps=mean(steps))
weekendPlot

weekdayPlot <- fullData2 %>%
    filter(weekend==0) %>%
    group_by(interval) %>%
    summarise(steps=mean(steps))
weekdayPlot

par(mfrow = c(2,1))
plot(weekdayPlot$interval,weekdayPlot$steps,type="l",xlab="Interval",ylab="Avg Steps", main = "Weekdays")
plot(weekendPlot$interval,weekendPlot$steps,type="l",xlab="Interval",ylab="Avg Steps", main = "Weekends")
```



